---
title: "patreco-week-4"
author: "wildhtob"
date: "7 5 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises Week 4

## Libraries

```{r libraries, message=FALSE, warning=FALSE}

library(readr)        
library(dplyr)        
library(ggplot2)      
library(lubridate)
library(sf) 

```

## Functions

Trying out some dummy functions. Normally, all the functions would go here but the actual functions for this exercise are to be found in Task 1.

```{r dummy functions, message=FALSE}

# simple funtion to demonstrate its structure
testfun <- function() {print("this function does nothing")}
testfun()

# calculcating age based on birthday
my_age <- function(birthday, units = "days") {
  difftime(Sys.time(), birthday, units = units)
}

my_age("1980-01-01", "weeks")



```

## Data Import

```{r data import, message=FALSE, warning=FALSE, include=TRUE}

wildschwein <- read_delim("data/wildschwein_BE_2056.csv", ",")


```

## Task 1: Writing a function

Creating a function for the Euclidian distance calculation. Two seperate functions. One for lead and one for lag. The default interval is set to 1 but can be changed, as we did in exercise 3 to calculate different steplengths.

```{r euclidian functions}

euclid_dist <- function(e1, e2, n1, n2) {
  sqrt((e1 - e2)^2 + (n1 - n2)^2)
}

euclid_lead <- function (df, interval = 1) {
  sqrt((df$E - lead(df$E, interval))^2 + 
       (df$N - lead(df$N, interval))^2)
}

euclid_lag <- function (df, interval = 1) {
  sqrt((lag(df$E, interval) - df$E)^2 + 
       (lag(df$N, interval) - df$N)^2)
}

```

Testing the function with the wildboar data.

```{r testing function, eval=FALSE, include=TRUE}

wildschwein$steplength1 <- euclid_lead(wildschwein)
wildschwein$steplength2 <- euclid_lead(wildschwein, 2)

```

## Task 2: Preparing analysis

Subset the wildboar data to simplify the analysis. Only use the individuals *Rosa* and *Sabi* for the timespan *01.04.2015* to *15.05.2015*.

```{r subsetting data, message=FALSE, warning=FALSE}

str(wildschwein) #Datetime is already formated correctly

time_start <- as.POSIXct("2015-04-01", tz = "UTC") # UTC is the default timzone
time_end <- as.POSIXct("2015-04-16", TZ = "UTC")

wildschwein_subset <- wildschwein %>% 
  filter(
    DatetimeUTC >= time_start & DatetimeUTC <= time_end,
    TierName == "Sabi" | TierName== "Rosa"
    )

```

## Taks 3: Create Join Key

Round the time stamps in order to compare Rosa and Sabis' locations and to be able to apply a join. Round the minutes of `DatetimeUTC` to a multiple of 15 and put the values in a new column by using `round_date()`

```{r round datetime}

wildschwein_subset$time_rounded <- wildschwein_subset$DatetimeUTC %>% 
  round_date("15 minutes")

head(wildschwein_subset)

```

## Task 4: Measuring distance at concurrent locations

Splitting the subset data into one df per animal.

```{r split dataset}

wildschwein_split <- wildschwein_subset %>% 
  group_split(TierName) %>% 
  setNames(c("Rosa", "Sabi")
  )

list2env(wildschwein_split, envir = .GlobalEnv)


```

Rejoining the dataframes by `time_rounded` using `inner_join` because only data points with temporal closeness are interesting here.

```{r inner join}


wildschwein_join <- inner_join(
  Sabi, Rosa, by = "time_rounded", suffix = c("_sabi", "_rosa")
  )


```

Calculating the Euclidian distance using the definied function `euclid_dist`

```{r calculating euclidian distance}
wildschwein_join <- wildschwein_join %>%  
  mutate(
    distance = euclid_dist(.$E_sabi, .$E_rosa, .$N_sabi, .$N_rosa)
    )
```

Or manually:

```{r calculating euclidian distance (no fn), eval=FALSE, include=TRUE}
wildschwein_join <- wildschwein_join %>%
    mutate(
    distance = sqrt((E_sabi - E_rosa)^2 + (N_sabi - N_rosa)^2)
    )
```

Calculating spatial closeness using a distance threshold.

```{r caluclating meet}

wildschwein_join$meet <- ifelse(wildschwein_join$distance <= 100, TRUE, FALSE)

```

## Task 5: Visualise data

```{r ggplot task 5}

ggplot(filter(wildschwein_join, meet == TRUE)) +
  geom_point(data = wildschwein_subset, aes(E, N, colour = TierName), alpha = 0.5) +
  geom_point(aes(E_rosa, N_rosa),
    shape = 21, colour = "black",
    fill = "#F8766D", stroke = 1, size = 2, alpha = 1
  ) +
  geom_point(aes(E_sabi, N_sabi),
    shape = 21, colour = "black",
    fill = "#00BFC4", stroke = 1, size = 2, alpha = 1
  ) +
  xlim(2570000, 2571000) +
  ylim(1204500, 1205500)
  
# The legend is missing the meeting points. To be resolved.

```

